sdmTMB: Fast, flexible, and user-friendly spatial and spatiotemporal generalized linear mixed-effects models

### 2021-11-18 doc/paper.Rmd
Develop paper

## Quantities of interest

@@

@@

## Advanced options

While not of interest to all users, sdmTMB also implements several advanced options for model fitting. 
First, models can include penalized likelihoods via assigning priors to model parameters.
These priors may be useful in cases where estimation is difficult because of identifiability issues or relatively flat likelihood surfaces.
Following other recent SPDE approaches [@simpson2017; @osgood-zimmerman2021], 'penalized complexity' priors may be particularly useful in assigning priors to the spatial range and variance.
Second, we allow non-linear effects of covariates to be modeled with penalized regression splines ("P-splines") [@eilers1996], in one or multiple dimensions, using the same `s(x)` syntax familiar to users of `mgcv`.
Third, for users interested in Bayesian inference, we allow models to be passed to Stan (cite) via the package tmbstan [@monnahan2018].

@@

An sdmTMB model requires a data frame that contains a response column, columns for any predictors, and columns for spatial coordinates.
It usually makes sense to convert the spatial coordinates to UTMs such that distance remains constant throughout the study region (e.g., using `sf::st_transform()`).
Here, we illustrate a spatial model fit to Pacific Cod (*Gadus macrocephalus*) trawl survey data from Queen Charlotte Sound, BC, Canada.
Our model contains a main effect of depth as a penalized smoother, a spatial random field, and Tweedie observation error.
Our data frame `pcod` (built into the package) has a column `year` for the year of the survey, `density` for density of Pacific Cod in a given survey tow, `depth` for depth in meters of that tow, and spatial coordinates `X` and `Y`, which are UTM coordinates in kilometers:

@@

We can make predictions with the `predict()` method; we will predict on new data with the `newdata` argument.
The data frame `qcs_grid` is a grid that extends over the full survey domain and contains all predictors used in the model set at values for which we we want to predict.
The output of `predict()` is a data frame containing overall estimates, estimates from the non-random-field components (intercept and depth), and estimates from the random field components.
We show a basic plot of the estimated spatial random field, predictions across a depth gradient, and predictions in space in Fig. \@ref(fig:pcod-fig)b--d.

@@

We could extend our model to a spatiotemporal model simply by supplying the year column name to the `time` argument and specifying how we want the spatiotemporal random fields to be structured.
Here we will turn off the spatial field and structure the spatiotemporal random fields as AR(1):

>>>>

@@

## Parameter priors/penalties

sdmTMB models can include penalized likelihoods by assigning priors to model parameters (`?sdmTMBpriors`).
These priors may be useful in cases where estimation is difficult because of identifiability issues, relatively flat likelihood surfaces, to impart prior information, or to achieve regularization.
Following other recent SPDE implementations [@simpson2017; @osgood-zimmerman2021], penalized complexity priors (`pc_matern`) may be particularly useful to constrain the spatial range and variance parameters.

## Derived quantities of interest

@@

<!-- ## Advanced options -->

<!-- While not of interest to all users, sdmTMB also implements several advanced options for model fitting.  -->
<!-- First, models can include penalized likelihoods via assigning priors to model parameters. -->
<!-- These priors may be useful in cases where estimation is difficult because of identifiability issues or relatively flat likelihood surfaces. -->
<!-- Following other recent SPDE approaches [@simpson2017; @osgood-zimmerman2021], 'penalized complexity' priors may be particularly useful in assigning priors to the spatial range and variance. -->
<!-- Second, we allow non-linear effects of covariates to be modeled with penalized regression splines ("P-splines") [@eilers1996], in one or multiple dimensions, using the same `s(x)` syntax familiar to users of `mgcv`. -->
<!-- Third, for users interested in Bayesian inference, we allow models to be passed to Stan (cite) via the package tmbstan [@monnahan2018]. -->

@@

An sdmTMB model requires a data frame that contains a response column, columns for any predictors, and columns for spatial coordinates.
It usually makes sense to convert the spatial coordinates to UTMs such that distance remains constant throughout the study region (e.g., using `sf::st_transform()`).
Here, we illustrate a spatial model fit to Pacific cod (*Gadus macrocephalus*) trawl survey data from Queen Charlotte Sound, BC, Canada.
Our model contains a main effect of depth as a penalized smoother, a spatial random field, and Tweedie observation error.
Our data frame `pcod` (built into the package) has a column `year` for the year of the survey, `density` for density of Pacific Cod in a given survey tow, `depth` for depth in meters of that tow, and spatial coordinates `X` and `Y`, which are UTM coordinates in kilometers:

@@

We can make predictions with the `predict()` method; we will predict on new data with the `newdata` argument.
The data frame `qcs_grid` (included in the package data) is a grid that extends over the full survey domain and contains all predictors used in the model set at values for which we we want to predict.
The output of `predict()` is a data frame containing overall estimates, estimates from the non-random-field components (intercept and depth), and estimates from the random field components.
We show a basic plot of the estimated spatial random field, predictions across a depth gradient, and predictions in space in Fig. \@ref(fig:pcod-fig)b--d.

@@

We could extend our model to a spatiotemporal one simply by supplying the year column name to the `time` argument and specifying how we want the spatiotemporal random fields to be structured.
Here we will turn off the spatial field and structure the spatiotemporal random fields as AR(1):
